---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "man/figures/"
)
```

# aricode: a package for efficient computations of standard clustering comparison measures

 <!-- badges: start -->
[![R build status](https://github.com/jchiquet/aricode/workflows/R-CMD-check/badge.svg)](https://github.com/jchiquet/aricode/actions)
[![CRAN Status](https://www.r-pkg.org/badges/version/aricode)](https://CRAN.R-project.org/package=aricode)
[![Coverage status](https://codecov.io/gh/jchiquet/aricode/branch/master/graph/badge.svg)](https://codecov.io/gh/jchiquet/aricode)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-blue.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![](https://img.shields.io/github/last-commit/jchiquet/aricode.svg)](https://github.com/jchiquet/aricode/commits/master)
  <!-- badges: end -->

## Installation

Stable version on the [CRAN](https://cran.rstudio.com/web/packages/aricode/).

```{r install_cran, eval = FALSE}
install.packages("aricode")
```

The development version is available via:

```{r install_github, eval = FALSE}
devtools::install_github("jchiquet/aricode")
```

## Description

Traditional implementations (e.g., function `adjustedRandIndex` of package **mclust**) are in $\Omega(n + u v)$ where 

- $n$ is the size of the vectors the classifications of which are to be compared,
- $u$ and $v$ are the respective number of classes in each vectors. 

Here, the implementation is in $\Theta(n)$, plus the gain of speed due to the `C` implementation.

Available measures are describes in the paper of Vinh et al, JMLR, 2009 (see reference below).

Vinh, Nguyen Xuan, Julien Epps, and James Bailey. *"Information theoretic measures for clusterings comparison: Variants, properties, normalization and correction for chance."* Journal of Machine Learning Research 11.Oct (2010): 2837-2854.

A modified adjusted Rand Index (unbiased and motivated by a multinomial model of the clusterings) is available.

A fast implementation for computing the Chi-square statistics is also implemented.


## Available functions

The functions included in aricode are:

- `ARI`: computes the adjusted rand index
- `Chi2`: computes the Chi-square statistics
- `MARI`: computes the modified adjusted rand index (Sundqvist et al, in preparation)
- `NVI`: computes the the normalized variation information
- `NID`: computes the normalized information distance
- `NMI`: computes the normalized mutual information
- `AMI`: computes the adjusted mutual information
- `entropy`: computes the conditional and joint entropies
- `clustComp`: computes all clustering comparison measures at once

## Timings

Here are some timings to compare the cost of computing the adjusted Rand Index with **aricode** or with the commonly used function `adjustedRandIndex` of the *mclust* package: the cost of the latter can be prohibitive for large vectors: 

```{r timings_function, echo=FALSE, message=FALSE, warning=FALSE}
library(aricode)
library(mclust)
library(ggplot2)

time.aricode <- function(times, c1, c2){
  replicate(times, system.time(ARI(c1, c2))[3])
}

time.mclust <- function(times, c1, c2){
  replicate(times, system.time(mclust::adjustedRandIndex(c1, c2))[3])
}

time.method <- function(times, c1, c2, n){
  rbind(
    data.frame(time = time.aricode(times, c1, c2), expr = "aricode", n = n),
    data.frame(time = time.mclust(times, c1, c2), expr = "mclust", n = n)
  )
}

# with similar classif, number of classes grows with n
sim.timings <- function(n, times = 10) {
    c1 <- sample(1:(n/200), n, replace=TRUE);c2 <- c1;
    i_change <- sample(1:n, n/50, replace=FALSE)
    c2[i_change] <- c2[rev(i_change)]
    out <- time.method(times, c1, c2, n)
    data.frame(time=out$time, method=out$expr, n = n)
}
```

```{r timings_run, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
# with similar classif, number of classes grows with n
ns <- sort(c(200 * 2^(3:14), 150 * 2^(3:15)))
timings <- do.call("rbind", lapply(ns, sim.timings))
```

```{r timings_plot, echo=FALSE, message=FALSE, warning=FALSE}
p.timings <- ggplot(timings, aes(x=n, y=time, colour=method)) +
  geom_smooth(data = dplyr::filter(timings, n > 1e4), method = "lm") + geom_point(size=0.25, alpha=0.9) + labs(y="time (sec.)") +
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
 scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   annotation_logticks()                 

p.timings + ggtitle("number of classes grows with n") + theme_bw()
```

