```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# aricode: a package for efficient computations of standard clustering comparison measures

## Installation

```
devtools::install_github("jchiquet/aricode")
```

## Description

Traditional implementations (e.g., function `adjustedRandIndex` of package **mclust**) are in $\Omega(n + u v)$ where 

- $n$ is the size of the vectors the classifications of which are to be compared,
- $u$ and $v$ are the respective number of classes in each vectors. 

Here, the implementation is in $\Theta(n)$, plus the gain of speed due to the `C` implementation.

Available measures are describes in the paper of Vinh et al, JMLR, 2009 (see reference below).

Vinh, Nguyen Xuan, Julien Epps, and James Bailey. *"Information theoretic measures for clusterings comparison: Variants, properties, normalization and correction for chance."* Journal of Machine Learning Research 11.Oct (2010): 2837-2854.

## Available functions

The functions included in aricode are:

- `ARI`: computes the adjusted rand index
- `NVI`: computes the the normalized variation information
- `NID`: computes the normalized information distance
- `NMI`: computes the normalized mutual information
- `entropy`: computes the conditional and joint entropies
- `clustComp`: computes all clustering comparison measures at once

## Timings

```{r timings_function, echo=FALSE}
library(aricode)
library(mclust)
library(ggplot2)

time.aricode <- function(times, c1, c2){
  replicate(times, system.time(ARI(c1, c2))[3])
}

time.mclust <- function(times, c1, c2){
  replicate(times, system.time(mclust::adjustedRandIndex(c1, c2))[3])
}

time.method <- function(times, c1, c2, n){
  rbind(
    data.frame(time = time.aricode(times, c1, c2), expr="aricode", n=n),
    data.frame(time = time.mclust(times, c1, c2), expr="mclust", n=n)
  )
}

# with random classif, number of classes grows with sqrt(n)
sim.timings.a <- function(n, times=20) {
    cat("",n)
    c1 <- sample(1:(sqrt(n)/2), n, replace=TRUE)
    c2 <- sample(1:(sqrt(n)/2), n, replace=TRUE)
    out <- time.method(times, c1, c2, n)
    return(data.frame(time=out$time, method=out$expr, n = n))
}

# with random classif, constant number of classes
sim.timings.b <- function(n, times=20) {
    cat("",n)
    c1 <- sample(1:100, n, replace=TRUE)
    c2 <- sample(1:100, n, replace=TRUE)
    out <- time.method(times, c1, c2, n)
    return(data.frame(time=out$time, method=out$expr, n = n))
}

# with similar classif, number of classes grows with n
sim.timings.c <- function(n, times=20) {
    cat("",n)
    c1 <- sample(1:(n/200), n, replace=TRUE);c2 <- c1;
    i_change <- sample(1:n, n/50, replace=FALSE)
    c2[i_change] <- c2[rev(i_change)]
    out <- time.method(times, c1, c2, n)
    return(data.frame(time=out$time, method=out$expr, n = n))
}
```

```{r timings_run, cache=TRUE, echo=FALSE}
ns <- sort(c(200 * 2^(3:14), 150 *2^(3:14)))

##  with random classif and nb of classes growing with n
timings.a <- do.call("rbind", lapply(ns, sim.timings.a))

##  with random classif and nb of classes growing with n
timings.b <- do.call("rbind", lapply(ns, sim.timings.b))

## with random classif and cst nb of classes (should be linear)
timings.c <- do.call("rbind", lapply(ns, sim.timings.c))
```

```{r timings_plot, echo=FALSE}
p.timings1 <- ggplot(timings.1, aes(x=n, y=time*scale.time, colour=method)) +
    geom_smooth() + geom_point(size=0.25, alpha=0.9) + labs(y="time (sec.)") +
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
 scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   annotation_logticks()                 
    
p.timings1a <- ggplot(timings.1a, aes(x=n, y=time*scale.time, colour=method)) +
    geom_smooth() + geom_point(size=0.25, alpha=0.9) + labs(y="time (sec.)") +
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
 scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   annotation_logticks()        

p.timings1b <- ggplot(timings.1b, aes(x=n, y=time*scale.time, colour=method)) +
  geom_smooth() + geom_point(size=0.25, alpha=0.9) + labs(y="time (sec.)") +
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
 ) +
 scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) +
   annotation_logticks()                 

grid.arrange(
    p.timings1 + ggtitle("Random classification, # of classes growth with n") + theme(legend.position="none",plot.margin= unit(c(0,0.05,0,0), "lines")),
    p.timings1a + ggtitle("Random classification, # of classes growth with sqrt(n)") + theme(legend.position="none",plot.margin= unit(c(0,0.05,0,0), "lines")),
    p.timings1b + ggtitle("Random classification, 100 classes") + theme(legend.position="none",plot.margin= unit(c(0,0.05,0,0), "lines")),
    g_legend(p.timings1), top="", nrow=1, widths=c(10,10,10,2)
)
```

